---
title: "first_big_data_project"
author: "Ariel"
date: "2024-07-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load libraries
```{r libraries}
setwd("C:/Users/Acer/big_data")
library(tidyverse)
library(ggplot2)
library(dplyr)
library(readr)
# Load necessary packages

```

```{r setup, include=FALSE}
library(pacman)
p_load(tidyverse, readr, skimr, knitr, ggplot2)
```


## Part 1: Descriptive Statistics

### 1.1 Load Data and Libraries
###First, load the necessary libraries and data.
```{r}
listings_data <- read_csv('listings_clean.csv')
calendar_data <- read_csv('calendar_clean.csv')
```
### 1.2 Descriptive Statistics
### Use skimr to provide a summary of the dataset.

```{r descriptive-statistics}
skim(listings_data)
```
## Part 2: Graphs for Single Variables
### 2.1 Graph 1: Distribution of price_dollars
* Price Distribution helps you understand the pricing landscape and adjust strategies accordingly.

* Why It’s Interesting:

1. Understand Pricing Trends: This graph shows how prices are distributed across all listings. You can identify common price ranges and see if there are price clusters or outliers.

2. Market Positioning: By understanding the distribution, you can determine whether most listings are priced higher or lower than average, which helps in positioning your listings competitively.

3. Revenue Potential: If you see that most listings are clustered in a specific price range, you might consider adjusting your pricing strategy to align with these clusters or to differentiate your listings.

4. Identify Outliers: Extreme prices, either very high or very low, can be investigated further. High prices might indicate premium listings, while low prices could be underpriced relative to their value.
```{r price-distribution}
ggplot(listings_data, aes(x = price_dollars)) +
  geom_histogram(binwidth = 10, fill = "blue", color = "black") +
  labs(title = "Distribution of Prices",
       x = "Price in Dollars",
       y = "Frequency")

```
Interpretation of the Price Distribution Graph
The graph you generated shows the distribution of listing prices in dollars. Here are some key observations and their implications:

Skewed Distribution:

The distribution is highly skewed to the right, with a large concentration of listings priced below $500. There are very few listings with prices above $1000, and those above $2000 are even rarer.
Implication: This skewness indicates that most listings are relatively affordable, while high-end, expensive listings are much less common. Airbnb might consider focusing on marketing these high-end listings differently to attract the right clientele.
Outliers:

The presence of prices that go up to $4000 suggests outliers in the data. These could be luxury properties or special cases.
Implication: Identifying these outliers can be useful for creating targeted marketing strategies. For example, Airbnb can create exclusive campaigns highlighting these luxury properties.
High Frequency of Low Prices:

The high frequency of listings priced below $500 suggests that Airbnb's market in this region is more inclined towards budget-friendly or mid-range options.
Implication: Airbnb could leverage this insight to enhance budget-friendly amenities or introduce more competitive pricing strategies to attract more budget-conscious travelers.



### 2.2 Graph 2: Reviews per Month
* Reviews Per Month indicates popularity and engagement, helping you target growth areas.

* Why It's Interesting:

1. Listing Popularity: This graph shows how frequently listings are reviewed, which is a good indicator of their popularity and booking frequency. More reviews typically correlate with higher demand.

2. Engagement Level: Frequent reviews may indicate that listings are in high demand or that they are well-managed and attract a lot of guest attention.

3. Seasonal Trends: If you have temporal data, you can analyze whether the number of reviews changes with seasons or events, helping to forecast demand and adjust strategies accordingly.

4. Opportunity for Growth: Listings with fewer reviews might have potential for growth. Strategies could be developed to increase their visibility or improve their appeal.

```{r reviews-per-month}
ggplot(listings, aes(x = reviews_per_month)) +
  geom_histogram(binwidth = 0.5, fill = "green", color = "black") +
  labs(title = "Distribution of Reviews per Month",
       x = "Reviews per Month",
       y = "Frequency")
```
Interpretation of the Reviews per Month Distribution Graph
The graph displays the distribution of the average number of reviews per month for the listings. Here are some key observations and their implications:

Skewed Distribution:

The distribution is highly skewed to the right, with the majority of listings receiving between 0 and 2 reviews per month.
Implication: Most listings are not receiving a high volume of reviews, indicating either low booking rates or less engagement from guests in leaving reviews. Airbnb might consider incentivizing guests to leave reviews to increase engagement.
High Frequency of Low Reviews:

A significant number of listings receive very few reviews (0-2 reviews per month). This could be due to new listings, less popular locations, or lower overall bookings.
Implication: Listings with few reviews may benefit from promotional activities to increase visibility and bookings. Additionally, Airbnb can introduce programs to encourage guests to leave reviews.
Presence of Listings with Higher Reviews:

There are listings that receive up to 20 reviews per month, although these are much less common.
Implication: These high-review listings could be top-performing properties, likely offering exceptional service or amenities. Understanding what makes these listings successful could help improve other listings.


### 2.3 Graph 3: Distribution of review_scores_rating
* Review Scores Rating gives insights into customer satisfaction and quality standards.

* Why It's Interesting:

1. Customer Satisfaction: This graph helps you see how review ratings are distributed. Higher ratings generally reflect better guest experiences and satisfaction.

2. Quality Indicators: Listings with higher ratings are often of higher quality or provide better service. Understanding this distribution can help you identify what is considered a “good” rating and set performance benchmarks.

3. Performance Benchmarking: If most listings have high ratings, your goal might be to maintain or exceed these standards. Conversely, if ratings are lower, there could be opportunities to improve quality or service.

4. Market Perception: Review scores can also impact how listings are perceived by potential guests. Higher-rated listings might attract more bookings and allow for higher pricing.
```{r rating-distribution}
ggplot(listings, aes(x = review_scores_rating)) +
  geom_histogram(binwidth = 5, fill = "red", color = "black") +
  labs(title = "Distribution of Review Scores Rating",
       x = "Review Scores Rating",
       y = "Frequency")

```


## Part 3: Graphs for Relationships Between Variables
### 3.1 Graph 4: Price vs. Reviews per Month
?????????? ?????????? ???????? ?????????? ???????? ???? ???? ???????? ?????????? ?????? ????????
```{r price-vs-reviews}
ggplot(listings, aes(x = price_dollars, y = reviews_per_month)) +
  geom_point() +
  labs(title = "Price vs. Reviews per Month",
       x = "Price in Dollars",
       y = "Reviews per Month")
```

### 3.2 Graph 5: Price vs. Rating
```{r price-vs-rating}
ggplot(listings, aes(x = price_dollars, y = review_scores_rating)) +
  geom_point() +
  labs(title = "Price vs. Review Scores Rating",
       x = "Price in Dollars",
       y = "Review Scores Rating")
```
### 3.3 Graph 6: Beds vs. Price
```{r beds-vs-price}
ggplot(listings, aes(x = beds, y = price_dollars)) +
  geom_boxplot() +
  labs(title = "Beds vs. Price",
       x = "Number of Beds",
       y = "Price in Dollars")
``` 

### 3.4 Graph 6: rating vs. accuracy
If the scatter plot shows a positive correlation (i.e., as the review_scores_accuracy increases, the review_scores_rating also increases), it suggests that guests tend to give higher overall ratings when the description of the listing is accurate.

```{r rating-vs-accuracy}
ggplot(listings, aes(x = review_scores_accuracy, y = review_scores_rating)) +
  geom_point() +
  geom_smooth(method = "lm", col = "blue") +
  labs(title = "Review Scores Rating vs. Review Scores Accuracy",
       x = "Review Scores Accuracy",
       y = "Review Scores Rating") +
  theme_minimal()
```



